---
title: "Time Series Analysis of Case Load, Recovery and Mortality Rates"
author: "Usman Khaliq"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(gganimate)
library(lubridate)

# Parameters

#Paths for the raw data files
covid_confirmed_data <- here::here("data/covid_confirmed.rds")
covid_deaths_data <- here::here("data/covid_deaths.rds")
covid_recovered_data <- here::here("data/covid_recovered.rds")
#==========================================================================

# Code
```

## Time series analysis on Confirmed COVID cases over time in China

First, lets read all the data into Rds objects
```{r}
covid_confirmed <-
  covid_confirmed_data %>% 
  read_rds() %>% 
  rename(
    province_state = `Province/State`,
    country_region = `Country/Region`,
    lat = Lat,
    long = Long
  )

covid_deaths <-
  covid_deaths_data %>% 
  read_rds() %>% 
  rename(
    province_state = `Province/State`,
    country_region = `Country/Region`,
    lat = Lat,
    long = Long
  )

covid_recovered <-
  covid_recovered_data %>% 
  read_rds() %>% 
  rename(
    province_state = `Province/State`,
    country_region = `Country/Region`,
    lat = Lat,
    long = Long
  )
```

Now, lets wrangle the data for the confirmed cases so that we can see the change in the values of confirmed cases over time.
(note - currently, look at country figures) 

```{r}

covid_confirmed %>% 
  pivot_longer(
    cols = -c(province_state, country_region, lat, long),
    names_to = "date",
    values_to = "confirmed_cases"
  ) %>% 
  group_by(country_region, date) %>% 
  summarise(confirmed_cases = sum(confirmed_cases)) %>% 
  mutate(date = mdy(date)) %>% 
  filter(country_region == "China") %>% 
  ggplot(aes(date, confirmed_cases)) +
  geom_line() +
  transition_reveal(date) +
  scale_x_date(
    date_labels = "%d %b %Y",
    date_breaks = "1 week"
  ) +
  labs(
    title = "Trajectory of Confirmed COVID-19 Cases in China",
    x = "Date",
    y = "Number of Cases",
    caption = "Source: Johns Hopkins University"
  )

```

## Time series analysis on COVID Deaths cases over time in China

```{r}
covid_deaths %>% 
  pivot_longer(
    cols = -c(province_state, country_region, lat, long),
    names_to = "date",
    values_to = "confirmed_deaths"
  ) %>% 
  group_by(country_region, date) %>% 
  summarise(confirmed_deaths = sum(confirmed_deaths)) %>% 
  mutate(date = mdy(date)) %>% 
  filter(country_region == "China") %>% 
  ggplot(aes(date, confirmed_deaths)) +
  geom_line() +
  transition_reveal(date) +
  scale_x_date(
    date_labels = "%d %b %Y",
    date_breaks = "1 week"
  ) +
  labs(
    title = "Trajectory of Confirmed COVID-19 Deaths in China",
    x = "Date",
    y = "Number of Deaths",
    caption = "Source: Johns Hopkins University"
  )
```

## Time series analysis on COVID Recoveries over time in China

```{r}
covid_recovered %>% 
  pivot_longer(
    cols = -c(province_state, country_region, lat, long),
    names_to = "date",
    values_to = "confirmed_recovered"
  ) %>% 
  group_by(country_region, date) %>% 
  summarise(confirmed_recovered = sum(confirmed_recovered)) %>% 
  mutate(date = mdy(date)) %>% 
  filter(country_region == "China") %>% 
  ggplot(aes(date, confirmed_recovered)) +
  geom_line() +
  transition_reveal(date) +
  scale_x_date(
    date_labels = "%d %b %Y",
    date_breaks = "1 week"
  ) +
  labs(
    title = "Trajectory of Confirmed COVID-19 Recoveries in China",
    x = "Date",
    y = "Number of Rexoveries",
    caption = "Source: Johns Hopkins University"
  )
```

